Type cones of $\b{g}$-vector fans

### 2019-02-11 thirdPaper/draft/typeConeAssociahedra.tex
More with Section 3

\subsection{Statement of the theorem}
We need to introduce some notations before stating the main result of this section.  Let~$\Lambda:= \End{\cat}(T)$, and let~$F$ be the functor
\[
 F=\cat(T,-):\cat \xrightarrow{} \MOD \Lambda.
\]
\begin{proposition}[\cite{BuanMarshReiten, KellerReiten}]
 The functor~$F$ induces an equivalence of~$\field$-linear categories
 \[
  F:\cat/(\susp T) \xrightarrow{} \MOD \Lambda,
 \]
 where~$(\susp T)$ is the ideal of morphisms factoring through an object of~$\add(\susp T)$ and~$\MOD \Lambda$ is the category of finite-dimensional right~$\Lambda$-modules.  This equivalence induces further equivalences between~$\add(T)$ and the category of projective modules, and between~$\add(\susp^2 T)$ and that of injective modules.
\end{proposition}

@@

\begin{proposition}[]\pg{Cite Happel}
 Any triangulated category admitting a Serre functor has almost-split triangles.  In particular, any~$2$-Calabi--Yau triangulated category has almost-split triangles.
\end{proposition}

@@

We can finally state the main theorem of this section. \pg{recall Auslander}

@@

 \item If~$B \notin \add(\susp T)$, then
 \[
  \langle A, r_B \rangle = \begin{cases}
                               0 & \textrm{if~$A\ncong B$;} \\
                               \dim_{\field}\field_B & \textrm{if~$A\cong B$.}
                              \end{cases}
 \]
\end{enumerate} 
\end{lemma}
\begin{proof}
 We only prove the first assertion; the second one is proved dually.  Assume that~$A\notin\add(\susp T)$.  Let
 \[
  A \xrightarrow{f} E \xrightarrow{} \susp^{-1} A \xrightarrow{h} \susp A
 \]
 be an almost-split triangle.  By Lemma~\ref{lemm::almost-split}, the morphism~$h$ is in~$(\susp T)$.  Applying the functor~$F=\cat(T,-)$, we get an exact sequence
 \[
  FA\xrightarrow{Ff} FE \xrightarrow{} F\susp^{-1} A \xrightarrow{} 0.
 \]
 Applying now the functor~$\Hom{\Lambda}(-, FB)$, we get an exact sequence
 \[
  0 \xrightarrow{} \Hom{\Lambda}(F\susp^{-1} A, FB) \xrightarrow{} \Hom{\Lambda}(FE, FB) \xrightarrow{Ff^*} \Hom{\Lambda}(FA, FB) \xrightarrow{} \coker(Ff^*) \xrightarrow{} 0
 \]
 If~$B\ncong A$, then the definition of an almost-split triangle implies that~$Ff^*$ is surjective. \pg{add an exact sequence here; image of Ffstar is non-invertible elements} Thus~$\coker(Ff^*) = 0$, and by additivity of the dimension in exact sequences, we get that~$\langle \ell_A, B \rangle =0$.

 If~$B\cong A$, then~$\coker(Ff^*)$ is isomorphic to the residue field of~$\End{\Lambda}(FA)$, which is isomorphic to the residue field~$\field_A$ of~$A$.
\end{proof}

>>>>

\subsection{Statement of the theorem}
We need to introduce some notations before stating the main result of this section.  Let~$\Lambda:= \End{\cat}(T)$, and let~$F$ be the functor
\[
 F=\cat(T,-):\cat \xrightarrow{} \MOD \Lambda.
\]
\begin{proposition}[\cite{BuanMarshReiten, KellerReiten}]\label{prop::functor-F}
 The functor~$F$ induces an equivalence of~$\field$-linear categories
 \[
  F:\cat/(\susp T) \xrightarrow{} \MOD \Lambda,
 \]
 where~$(\susp T)$ is the ideal of morphisms factoring through an object of~$\add(\susp T)$ and~$\MOD \Lambda$ is the category of finite-dimensional right~$\Lambda$-modules.  This equivalence induces further equivalences between~$\add(T)$ and the category of projective modules, and between~$\add(\susp^2 T)$ and that of injective modules.
\end{proposition}

@@

\begin{proposition}[Proposition I.2.3 of~\cite{ReitenVandenbergh}]
 Any triangulated category admitting a Serre functor has almost-split triangles.  In particular, any~$2$-Calabi--Yau triangulated category has almost-split triangles.
\end{proposition}

@@

We can finally state the main theorem of this section, which is an analogue of the main result of~\cite{Auslander1984}.

@@

 \item If~$B \notin \add(\susp T)$, then
 \[
  \langle A, r_B \rangle = \begin{cases}
                               0 & \textrm{if~$A\ncong B$;} \\
                               \dim_{\field}\field_B & \textrm{if~$A\cong B$.}
                              \end{cases}
 \]
\end{enumerate} 
\end{lemma}
\begin{proof}
 We only prove the first assertion; the second one is proved dually.  Assume that~$A\notin\add(\susp T)$.  Let
 \[
  A \xrightarrow{f} E \xrightarrow{} \susp^{-1} A \xrightarrow{h} \susp A
 \]
 be an almost-split triangle.  By Lemma~\ref{lemm::almost-split}, the morphism~$h$ is in~$(\susp T)$.  Applying the functor~$F=\cat(T,-)$, we get an exact sequence
 \[
  FA\xrightarrow{Ff} FE \xrightarrow{} F\susp^{-1} A \xrightarrow{} 0.
 \]
 Applying now the functors~$\cat(-, B)$ and~$\Hom{\Lambda}(-, FB)$, we get a commutative diagram whose rows are exact sequences and whose vertical maps are surjective by Proposition~\ref{prop::functor-F}.
 \[
  \xymatrix{ & \cat(\susp^{-1} A, B)\ar[r]\ar@{->>}[d] & \cat(E, B)\ar[r]^{f^*}\ar@{->>}[d] & \cat(A,B)\ar[r]\ar@{->>}[d] & \coker(f^*)\ar[r]\ar@{->>}[d] & 0 \\
  0 \ar[r] & \Hom{\Lambda}(F\susp^{-1} A, FB)\ar[r] & \Hom{\Lambda}(FE, FB)\ar[r]^{Ff^*} & \Hom{\Lambda}(FA, FB)\ar[r] & \coker(Ff^*)\ar[r] & 0.
  }
 \]

 If~$B\ncong A$, then the definition of an almost-split triangle implies that~$f^*$ is surjective.  Thus~$\coker(f^*)=0$, so that~$\coker(Ff^*) = 0$, and by additivity of the dimension in exact sequences, we get that~$\langle \ell_A, B \rangle =0$.

 If~$B\cong A$, then~$\coker(f^*)$ is isomorphic to the residue field~$\field_A$ of~$\cat(A,A)$.  Since~$A$ is not in~$\add(\susp T)$, the ideal of endomorphisms of~$A$ factoring through an object of~$\add(\susp T)$ is contained in the maximal ideal of~$\cat(A,A)$.  Therefore, the rightmost morphism in the commutative diagram is an isomorphism, so~$\coker(Ff^*)$ is isomorphic to~$\field_A$.
\end{proof}
